---
title: "MOFA model of aIg stimulated cells in prescence of ibrutinib"
output:
  workflowr::wflow_html:
    toc: TRUE
    code_folding: hide
editor_options:
  chunk_output_type: console
---


```{r setup, warning=FALSE, message=FALSE}
source("code/load_packages.R")

panellabels <- c('a', 'b', 'c','d' , 'e', 'f', 'g', 'h', 'i', 'j', 'k')

add.textsize <- theme(axis.text.x = element_text(colour = 'black', size = 7),
          axis.text.y = element_text(colour = 'black',size=7),
          text = element_text(size=7),
          axis.text=element_text(size=7),
          plot.title = element_text(size=7)
          )

colorgradient6 <- c("#d4d4d3","#859FCA", "#4D7CC6" ,"#1F5284","#11304C", "#0C2236" )
colorgradient7 <- c(colorgradient6,"orange2")
colorsibru <- c(colorgradient7[c(1,2)],"#E69F00", colorgradient7[c(6)], "#D55E00","#ff0000")


labels.withibru <- c(0, 6, "6 \n+ibru", 180, "180\n+ibru")
labels.withibru.selected <- c(0, 6, "6 \n+ibru", 180, "180\n+ibru","180**\n+ibru")


```

```{r seurat object}
seu_combined_selectsamples.withibru <- readRDS("output/seu_ibru_samples.rds")

proteins.all.withibru <- row.names(seu_combined_selectsamples.withibru[["PROT"]])

meta.allcells.withibru <- seu_combined_selectsamples.withibru@meta.data %>%
  mutate(sample = rownames(seu_combined_selectsamples.withibru@meta.data))

seu_combined_selectsamples.withibru <- FindVariableFeatures(seu_combined_selectsamples.withibru, selection.method = "mvp", assay = "SCT.RNA", num.bin =20, mean.cutoff = c(0, 5), dispersion.cutoff = c(0.5,10), nfeatures =500, verbose = FALSE)

genes.variable <- seu_combined_selectsamples.withibru@assays$SCT.RNA@var.features[-grep("^MT", seu_combined_selectsamples.withibru@assays$SCT.RNA@var.features)] # without the mitochondrial genes 

seu_combined_selectsamples.withibru
```

## MOFA analysis ibru data 

Note, if mofa object was generated before, it will read the generated rds file. (this will speed-up the process of generating this html document if edits are required)

```{r Do mofa analysis}
myfiles <- list.files(path="output/", pattern = ".rds$")

if("MOFA_ibru.rds" %in%  myfiles){
  
  mofa <- readRDS("output/MOFA_ibru.rds")} else { 

    ## Nested list of RNA and Protein data (retrieved from filtered Seurat object)    
     mofa <- create_mofa(list(
      "RNA" = as.matrix( seu_combined_selectsamples.withibru@assays$SCT.RNA@scale.data[genes.variable,] ),
      "PROT" = as.matrix( seu_combined_selectsamples.withibru@assays$PROT@scale.data[proteins.all.withibru,] ))
      )

     ## Default data, model and training options
     data_opts <- get_default_data_options(mofa)
     model_opts <- get_default_model_options(mofa)
     train_opts <- get_default_training_options(mofa)
     train_opts$seed <- 42 # use same seed for reproducibility
    
    mofa <- prepare_mofa(
      object = mofa,
      data_options = data_opts,
      model_options = model_opts,
      training_options = train_opts
      )

      mofa <- run_mofa(mofa, outfile = "output/MOFA_ibru.hdf5")
      mofa <- run_umap(mofa)
      samples_metadata(mofa) <- meta.allcells.withibru
      saveRDS(mofa, file= "output/MOFA_ibru.rds")
  
}

mofa
```




## Figure 3

### Main

```{r Figure panels, message=FALSE, warning=FALSE,echo = T, fig.show= 'hide'}
## UMAP
plot.umap.data <-  plot_dimred(mofa, method="UMAP", color_by = "condition",stroke = 0.001, dot_size =1, alpha = 0.2, return_data = T)

plot.umap.all <- ggplot(plot.umap.data, aes(x=x, y = y, fill = color_by))+
  geom_point(size = 0.7, alpha = 0.5, shape = 21, stroke = 0) +
  theme_half_open() +
  scale_fill_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg \n(minutes)")+
   theme(legend.position="none")+
  add.textsize +
  scale_x_reverse()+
  scale_y_reverse()+
  labs(title = "Ibrutinib affects signal transduction \nat minutes and hour time-scale", x = "UMAP 1", y = "UMAP 2")

## UMAP legend
legend.umap <- get_legend( ggplot(plot.umap.data, aes(x=x, y = y, fill = color_by))+
  geom_point(size = 2, alpha = 0.5, shape = 21, stroke = 0) +
  theme_half_open() +
  scale_fill_manual(values = colorsibru, labels = c(labels.withibru), name ="Time aIg \n(minutes)",)+
   theme(legend.position="right")+
  add.textsize +
  scale_x_reverse()+
  scale_y_reverse()+
  labs(title = "Ibrutinib affects signal transduction at \nminute and hour timepoints after aIG stimulation", x = "UMAP 1", y = "UMAP 2"))
legend.umap <- as_ggplot(legend.umap)

## correlation time and factors
plot.correlation.covariates.withibru <- correlate_factors_with_covariates(mofa, 
  covariates = c("time", "inhibitor"),
  factors = 8:1,
  plot = "r"
)

plot.correlation.covariates.withibru <- ggcorrplot(plot.correlation.covariates.withibru, tl.col = "black", method = "square", lab = TRUE, ggtheme = theme_void, colors = c("orange3", "white", "orange3"), lab_size = 2.5) +
  add.textsize +
  labs(title = "Correlation\ncoefficient\n", y = "") +
  scale_y_discrete(labels = c("time\naIg\n","ibrutinib\ntreatment\n")) +
  coord_flip() +
  theme(axis.text.x=element_text(angle =0,hjust = 0.5), 
        axis.text.y=element_text(size = 5),
        legend.position="none",
        plot.title = element_text(hjust = 0.5))

## functions violin prots
f.violin <- function(data, feature){
  
  ggplot(subset(data)  , aes(x = as.factor(condition), y =get(noquote(feature)))) +
    annotate("rect",
          xmin = 4 - 0.45,
             xmax = 5 + 0.5,
           ymin = -5.5, ymax =5, fill = "lightblue",
           alpha = .4
  )+
  geom_violin(alpha = 0.9,aes(fill = condition))+
  geom_jitter(width = 0.05,size = 0.1, color = "black")+ 
  stat_summary(fun=median, geom="point", shape=95, size=2, inherit.aes = T, position = position_dodge(width = 0.9), color = "red")+
  theme_few()+
  labs(title = paste0(feature)) +
  scale_x_discrete(labels = labels.withibru, expand = c(0.1,0.1), name = "Time aIg (Minutes)") +
  scale_y_continuous(expand = c(0,0), name = "Counts (scaled)") +
  scale_color_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",)+ 
  scale_fill_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",) +
  add.textsize +
  theme(axis.ticks=element_line(color="black", size = 0.2),
        legend.position="none") 
}

## functions violin RNA
f.violin.rectlarge <- function(data, feature){
  
  ggplot(subset(data)  , aes(x = as.factor(condition), y =get(noquote(feature)))) +
    annotate("rect",
          xmin = 4 - 0.45,
             xmax = 5 + 0.5,
           ymin = -3, ymax = 15, fill = "lightblue",
           alpha = .4
  )+
  geom_violin(alpha = 0.9,aes(fill = condition))+
  geom_jitter(width = 0.05,size = 0.1, color = "black")+ 
  stat_summary(fun=median, geom="point", shape=95, size=2, inherit.aes = T, position = position_dodge(width = 0.9), color = "red")+
  theme_few()+
  labs(title = paste0(feature)) +
  scale_x_discrete(labels = labels.withibru, expand = c(0.1,0.1), name = "Time aIg (Minutes)") +
  scale_y_continuous(expand = c(0,0), name = "Counts (scaled)") +
  scale_color_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",)+ 
  scale_fill_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",) +
  add.textsize +
  theme(axis.ticks=element_line(color="black", size = 0.2),
        legend.position="none") 
}

## functions violin factors
f.violin.fact <- function(data = proteindata.subset, protein){
  
  ggplot(subset(data)  , aes(x = as.factor(condition), y =get(noquote(protein)))) +
    annotate("rect",
          xmin = 4 - 0.5,
             xmax = 5 + 0.5,
           ymin = -4.5, ymax = 4.5, fill = "lightblue",
           alpha = .4
  )+
    geom_hline(yintercept=0, linetype='dotted', col = 'black')+ 
  geom_violin(alpha = 0.9,aes(fill = condition))+
  geom_jitter(width = 0.05,size = 0.1, color = "black")+ 
  stat_summary(fun=median, geom="point", shape=95, size=2, inherit.aes = T, position = position_dodge(width = 0.9), color = "red")+
  theme_few()+
  ylab(paste0(protein)) +
  scale_x_discrete(labels = labels.withibru, expand = c(0.1,0.1), name = "Time aIg (Minutes)") +
    scale_y_continuous(expand = c(0,0), name = strsplit(protein, split = "\\.")[[1]][2]) +
  scale_color_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",)+ 
  scale_fill_manual(values = colorsibru, labels = c(labels.withibru), name = "Time aIg (Minutes)",) +
  add.textsize +
  theme(legend.position="none") 
}


## Factor data for violins
MOFAfactors<- as.data.frame(mofa@expectations$Z) %>%
  mutate(sample = rownames(as.data.frame(mofa@expectations$Z)[,1:mofa@dimensions$K])) 

MOFAfactors <- left_join(as.data.frame(MOFAfactors), meta.allcells.withibru)

factors_toplot <- c(colnames(MOFAfactors)[c(1:mofa@dimensions$K)])

## plot violin factors
for(i in factors_toplot) {
assign(paste0("plot.violin.factor.", i), f.violin.fact(data = MOFAfactors,protein = i)) 
}

plot.violin.factor1 <- plot.violin.factor.group1.Factor1 +
  theme(#axis.title.x=element_blank(),
        #axis.text.x=element_blank(),
        #axis.ticks.x=element_blank(),
        legend.position="none") +
  scale_y_reverse(expand = c(0,0), name = "Factor value")+
  labs(title = "Factor 1 \n'BCR signaling'")

plot.violin.factor3 <- plot.violin.factor.group1.Factor3 +
  theme(axis.title.y=element_blank(),
        #axis.text.y=element_blank(),
        #axis.ticks.y=element_blank(),
        legend.position="none") +
  scale_y_continuous(expand = c(0,0), name = "Factor 3 value")+
  labs(title = "Factor 3 \n'B-cell activation'")


plot.violin.factor5 <- plot.violin.factor.group1.Factor5 +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position="none") +
  scale_y_continuous(expand = c(0,0), name = "Factor 5 value")+
  labs(title = "Factor 5 \n")

## Protein data for violins
proteindata <- as.data.frame(t(seu_combined_selectsamples.withibru@assays$PROT@scale.data)) %>%
  mutate(sample = rownames(t(seu_combined_selectsamples.withibru@assays$PROT@scale.data))) %>%
  left_join(meta.allcells.withibru, by = "sample")

proteinstoplot <- c("p-Erk1/2", "p-PLC-y2Y759", "p-BLNK", "p-CD79a", "p-Syk", "p-JAK1")

for(i in proteinstoplot) {
assign(paste0("plot.violin.prot.", i), f.violin(data = proteindata, feature = i))

}

`plot.violin.prot.p-Erk1/2` <- `plot.violin.prot.p-Erk1/2` +
  labs(title = "p-ERK1/2") 
  

## Data for violin genes
rnadata <- as.data.frame(t(seu_combined_selectsamples.withibru@assays$SCT.RNA@scale.data)) %>%
  mutate(sample = rownames(t(seu_combined_selectsamples.withibru@assays$SCT.RNA@scale.data))) %>%
  left_join(meta.allcells.withibru, by = "sample")

for(i in c("NEAT1", "NPM1", "BTF3", "RGS2", "RGS13","VEGFA")) {
assign(paste0("plot.violin.RNA.", i), f.violin.rectlarge(data = rnadata, feature = i))

}

## Enrichment analysis factor 5
weights.RNA <- get_weights(mofa, views = "RNA",as.data.frame = TRUE)

weights.RNA.filtered.f5 <- weights.RNA %>%
  mutate(Entrez = mapIds(org.Hs.eg.db, as.character(weights.RNA$feature), 'ENTREZID', 'SYMBOL'))  %>%
  filter(abs(value) >= 0.2 & factor == "Factor5") %>%
  mutate(sign = ifelse(value <= 0, "neg", "pos")) 

enriched.go.bp.fct5.clusterdposneg <- compareCluster(Entrez~factor+sign, data=weights.RNA.filtered.f5, fun='enrichGO', OrgDb='org.Hs.eg.db',ont= "BP",
                pAdjustMethod = "BH", readable = TRUE)

enriched.go.bp.fct5.clusterdposneg <- simplify(enriched.go.bp.fct5.clusterdposneg, cutoff=0.8, by="p.adjust", select_fun=min)


plot.enriched.go.f5.top5 <- dotplot(enriched.go.bp.fct5.clusterdposneg,x=~sign, showCategory = 5, by = "count") +
  scale_y_discrete(labels = c("negative regulation of G protein-coupled\nreceptor signaling pathway") , limits = "negative regulation of G protein-coupled receptor signaling pathway") +
  facet_grid(~factor) +
  add.textsize+ 
  scale_color_viridis(option="E", direction = -1) +
  scale_size_continuous(range=c(0.8, 2)) 

## Top enriched sets
topgeneset.fct5<- unlist(str_split(subset(enriched.go.bp.fct5.clusterdposneg@compareClusterResult, sign == "pos")$geneID, pattern = "/"))

topgeneset.fct5 = bitr(topgeneset.fct5, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

bottomgeneset.fct5<- unlist(str_split(subset(enriched.go.bp.fct5.clusterdposneg@compareClusterResult, sign == "neg")[1:5,]$geneID, pattern = "/"))

bottomgeneset.fct5 = bitr(bottomgeneset.fct5, fromType="SYMBOL", toType="ENTREZID", OrgDb="org.Hs.eg.db")

## PROT factor 5 loadings
plotdata.rank.PROT.5<-plot_weights(mofa, 
  view = "PROT", 
  factors = c(5), 
  nfeatures = 4, 
  text_size = 1,
  manual = list(c("p-Erk1/2", "XBP1_PROT"),NULL), 
  color_manual = list("black","black"),
  return_data = TRUE
)

plotdata.rank.PROT.5<- plotdata.rank.PROT.5%>%
  mutate(Rank = 1:nrow(plotdata.rank.PROT.5),
         Weight = value, 
         colorvalue = ifelse(labelling_group == 3,"black", ifelse(labelling_group == 2, "black", "grey2")),
         highlights = ifelse(labelling_group >= 1, as.character(feature), "")
         )

plot.rank.PROT.5<- ggplot(plotdata.rank.PROT.5, aes(x=Rank, y = Weight, label = highlights)) +
  labs(title = "<p><span style='color:black'></span>  phospho-proteins<span style='color:black'><span style='color:blue4'></span> ",  #
       x= "Loading rank\n",
       y= "Factor 5 loading value") +
  geom_point(size=0.1, color =plotdata.rank.PROT.5$colorvalue) +
  geom_text_repel(size = 2, 
                  segment.size = 0.2, 
                  color =plotdata.rank.PROT.5$colorvalue,
                  nudge_x       = -1 - plotdata.rank.PROT.5$Weight,
                  direction     = "y",
                  hjust         = 0,
                  segment.color = "grey50")+
  theme_few()+
  add.textsize +
  scale_x_continuous() +
  add.textsize +
  theme(
    plot.title = element_markdown(),
    axis.text.x=element_blank(),
    axis.ticks.x=element_blank(),
    axis.text.y = element_blank(),
    axis.title.y = element_blank(),
    axis.ticks.y = element_blank()
        )+
  ylim(c(-1,1))


### RNA loadings
plotdata.rank.RNA.5<-plot_weights(mofa, 
  view = "RNA", 
  factors = c(5), 
  nfeatures = 2, 
  text_size = 1,
  manual = list(topgeneset.fct5$SYMBOL,NULL), 
  color_manual = list("black","black"),
  return_data = TRUE
)

plotdata.rank.RNA.5<- plotdata.rank.RNA.5%>%
  mutate(Rank = 1:nrow(plotdata.rank.RNA.5),
         Weight = value, 
         colorvalue = ifelse(labelling_group == 3,"black", ifelse(labelling_group == 2, "black", "grey2")),
         highlights = ifelse(labelling_group >= 1, as.character(feature), "")
         )

plot.rank.RNA.5<- ggplot(plotdata.rank.RNA.5, aes(x=Rank, y = Weight, label = highlights)) +
  labs(title = "Top loadings  <p><span style='color:black'></span> genes <span style='color:black'><span style='color:black'></span> ",  #
       x= "Loading rank\n",
       y= "Factor 5 loading value") +
  geom_point(size=0.1, color =plotdata.rank.RNA.5$colorvalue) +
  geom_text_repel(size = 2, 
                  segment.size = 0.2, 
                  color =plotdata.rank.RNA.5$colorvalue,
                  nudge_x       = -1 - plotdata.rank.RNA.5$Weight,
                  direction     = "y",
                  hjust         =  0,
                  segment.color = "grey50")+
  theme_few()+
  add.textsize +
  scale_x_continuous() +
  add.textsize +
  theme(
    plot.title = element_markdown()
  )  +
  theme(axis.text.x=element_blank(),
        axis.ticks.x=element_blank()
        )+
  ylim(c(-1,1))


```


```{r Fig_3}
Fig3.row1 <- plot_grid(plot.umap.all, legend.umap, NULL, plot.correlation.covariates.withibru, plot.violin.factor1,plot.violin.factor3,plot.violin.factor5, labels = c(panellabels[1],"",panellabels[2], "", panellabels[3]), label_size = 10, ncol =7, rel_widths = c(1.2,0.1,0.1,0.55,0.72,0.68,0.61))

Fig3.row2 <- plot_grid(plot.rank.RNA.5, plot.rank.PROT.5,plot.violin.RNA.RGS2, plot.violin.RNA.RGS13, `plot.violin.prot.p-Erk1/2`, labels = c(panellabels[4], "", panellabels[5], "", panellabels[6]), label_size = 10, ncol = 5,  rel_widths = c(0.8,0.66,1.2,1.2,1.2))


plot.Fig3 <- plot_grid(Fig3.row1, Fig3.row2, labels = c( "", ""), label_size = 10, ncol = 1, rel_heights =c(0.9,0.7,1))

ggsave(plot.Fig3,filename = "output/paper_figures/Fig3.pdf", width = 183, height = 122, units = "mm",  dpi = 300, useDingbats = FALSE)
ggsave(plot.Fig3, filename = "output/paper_figures/Fig3.png", width = 183, height = 122, units = "mm",  dpi = 300)

plot.Fig3
```
*Figure 3.* aIG stimultion in contect of Ibrutinib.

### Supplementary MOFA model

```{r Suppl MOFAmodel panels, warning=FALSE,echo = T, fig.show= 'hide', message=FALSE}
## Variance per factor
plot.variance.perfactor.all <- plot_variance_explained(mofa, x="factor", y="view") +
    add.textsize +
    labs(title = "Variance explained by each factor per modality") 

## variance total
plot.variance.total <- plot_variance_explained(mofa, x="view", y="factor", plot_total = T)
plot.variance.total <- plot.variance.total[[2]] +
  add.textsize +
    labs(title = "Total \nvariance") +
  geom_text(aes(label=round(R2,1)), vjust=1.6, color="white", size=2.5)

## Significance correlation covariates
plot.heatmap.pval.covariates <- as.ggplot(correlate_factors_with_covariates(mofa, 
  covariates = c("time"),
  factors = 1:mofa@dimensions$K,
  fontsize = 7, 
  cluster_row = F,
  cluster_col = F
))+ 
  add.textsize +
  theme(axis.text.y=element_blank(),
        axis.text.x=element_blank(),
        plot.title = element_text(size=7, face = "plain"),
        )

## Factor values over time
plot.violin.factorall <- plot_factor(mofa, 
  factor = c(1:mofa@dimensions$K),
  color_by = "condition",
  dot_size = 0.2,      # change dot size
  dodge = T,           # dodge points with different colors
  legend = F,          # remove legend
  add_violin = T,      # add violin plots,
  violin_alpha = 0.9  # transparency of violin plots
) +   
  add.textsize +
  scale_color_manual(values=c(colorsibru, labels = c(labels), name = "Time aIg")) +
  scale_fill_manual(values=c(colorsibru, labels = c(labels), name = "Time aIg"))+
  labs(title = "Factor values per time-point of additional factors not correlating with time." ) +
  theme(axis.text.x=element_blank())

## Loadings factors stable protein

plot.rank.PROT.2.4to7 <- plot_weights(mofa, 
  view = "PROT", 
  factors = c(1:mofa@dimensions$K), 
  nfeatures = 3, 
  text_size = 1.5
) +   
  add.textsize +
  labs(title = "Top 3 Protein loadings per factor") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
)

## Loadings factors stable protein

plot.rank.RNA.2.4to7 <- plot_weights(mofa, 
  view = "RNA", 
  factors = c(1:mofa@dimensions$K), 
  nfeatures = 3, 
  text_size = 1.5
) +   
  add.textsize +
  labs(title = "Top 3 RNA loadings per factor") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()
)

## correlation time and factors
plot.correlation.covariates <- correlate_factors_with_covariates(mofa, 
  covariates = c("time"),
  factors = mofa@dimensions$K:1,
  plot = "r"
)

plot.correlation.covariates <- ggcorrplot(plot.correlation.covariates, tl.col = "black", method = "square", lab = TRUE, ggtheme = theme_void, colors = c("#11304C", "white", "#11304C"), lab_size = 2.5) +
  add.textsize +
  labs(title = "Correlation of \nfactors with \ntime of treatment", y = "") +
  scale_y_discrete(labels = "") +
  coord_flip() +
  theme(axis.text.x=element_text(angle =0,hjust = 0.5), 
        axis.text.y=element_text(size = 5),
        legend.position="none",
        plot.title = element_text(hjust = 0.5))
```


```{r Suppl_MOFA_model, fig.width=7.2, fig.height=8.66}
Fig.3.suppl.mofa.row1 <- plot_grid(plot.variance.perfactor.all, plot.variance.total,NULL, plot.heatmap.pval.covariates, labels = c(panellabels[1:3]), label_size = 10, ncol = 4, rel_widths = c(1.35, 0.3,0.25,0.38))

Fig.3.suppl.mofa.row2 <- plot_grid(plot.violin.factorall,legend.umap, labels = panellabels[4], label_size = 10, ncol = 2, rel_widths = c(1,0.1))



Suppl_mofa <- plot_grid(Fig.3.suppl.mofa.row1, Fig.3.suppl.mofa.row2, plot.rank.PROT.2.4to7, plot.rank.RNA.2.4to7, labels = c("","", panellabels[5:6]),label_size = 10, ncol = 1, rel_heights = c(1.45,1,1.1,1.1))


ggsave(Suppl_mofa, filename = "output/paper_figures/Fig3.Suppl_MOFAibru.pdf", width = 183, height = 220, units = "mm",  dpi = 300, useDingbats = FALSE)
ggsave(Suppl_mofa, filename = "output/paper_figures/Fig3.Suppl_MOFAibru.png", width = 183, height = 220, units = "mm",  dpi = 300)

Suppl_mofa
```
*Supplementary Figure.* MOFA model additional information

Enrichment genes factor 5 positve loadings (> 0.2).
```{r}
# print positive enrichment
subset(enriched.go.bp.fct5.clusterdposneg@compareClusterResult, sign == "pos")

```

### Supplementary violins

```{r Suppl_ibru_violins}
Suppl.row1 <- plot_grid(`plot.violin.prot.p-PLC-y2Y759`, `plot.violin.prot.p-BLNK`,`plot.violin.prot.p-CD79a`,`plot.violin.prot.p-Syk`,`plot.violin.prot.p-JAK1`,labels = panellabels[1], label_size = 10, ncol = 5, rel_widths = c(1,1,1,1,1))

Suppl.row2 <- plot_grid(plot.violin.RNA.NEAT1, plot.violin.RNA.NPM1, plot.violin.RNA.BTF3,labels = panellabels[2], label_size = 10, ncol = 5, rel_widths = c(1,1,1,1,1))

Suppl_ibru_protsgenes <- plot_grid(Suppl.row1, Suppl.row2,label_size = 10, ncol = 1, rel_heights = c(1,1))

ggsave(Suppl_ibru_protsgenes, filename = "output/paper_figures/Fig3.suppl.violinsibru.pdf", width = 183, height = 122, units = "mm",  dpi = 300, useDingbats = FALSE)
ggsave(Suppl_ibru_protsgenes, filename = "output/paper_figures/Fig3.suppl.violinsibru.png", width = 183, height = 122, units = "mm",  dpi = 300)

Suppl_ibru_protsgenes
```
*Supplementary Figure.* Violin plots of highlighted phopho-proteins & genes
